---
title: 'Sediment Stabilization experiments (2022 and 2023) on microbial mats'
output:
  html_document:
    df_print: paged
---

load packages:
```{r}
# load packages
library(ggplot2)
library(tidyr)
library(dplyr)
#working with time 
library(lubridate)
library(cowplot)
```


## Loading the data

The master excel sheet is all of my ooid data from any mat experiment tested, the sheet is different for each exp but has the same headers/information

ooid is always in ooid %

```{r}
#loading the combined xcel sheet 
may_23_dat <- readxl::read_xlsx("data/ooid_data_master.xlsx", sheet = "field_may_23")

feb_23_dat <- readxl::read_xlsx("data/ooid_data_master.xlsx", sheet = "field_feb_23")

july_23_dat <-  readxl::read_xlsx("data/ooid_data_master.xlsx", sheet = "field_july_23")

background_ooid <- readxl::read_excel("data/ooid_data_master.xlsx", sheet = "background")
```


#plot for ooids on mats through the mangrove highway 
```{r}
background_ooid$LatLong <- 
 paste(background_ooid$latlong, background_ooid$long, sep = ",")
```

#the actual Plot
```{r}
#avg is 8.82
background_ooid |>
#get ready to plot 
 ggplot()+
  #generate aes
  aes(x=`long`, y =`ooid %`, color = `location_num` ) +
#adding them as points
  geom_point(size = 3.5, show.legend = FALSE)+
  #adding a y intercept at the average 
  geom_hline(yintercept = 8.82, color='lightsalmon3') + 
  #adding text 
 annotate("text", x=-71.70265, 
          y=11, label="avg = 8.82%", 
          color="lightsalmon3") +
#labels
  labs(
   # this is left blank for the cowplot 
    x="", 
    y="ooid %", 
    title ='Natural ooid coverage on mats' ) +

#plot clean  
 theme( 
   legend.position = "right"
)+
  expand_limits(y=c(0, 80))+
  theme_bw()
```

Multi-day in situ field incubation experiment results (Feb, May, July 2023)

## Averaging the 3 data points for each experiment site
```{r}
#getting data into the average values
avg_may_dat <-may_23_dat|>
#grouping by the sample and the time
  group_by(sample, `hours after`)|>
#calculating the mean and the standard deviation from out 3 replicas at each site
  summarise(
    avg=mean(ooid), 
    sd = sd(ooid), 
    #keeping mat type to recall it later 
    mat_type = mat_type
  ) |>
#ungrouping
  ungroup()

```

```{r}
#getting data into the average values
avg_feb_dat <-feb_23_dat|>
#grouping by the sample and the time
  group_by(sample, `hours after`)|>
#calculating the mean and the standard deviation from out 3 replicas at each site
  summarise(
    avg=mean(ooid), 
    sd = sd(ooid), 
    #keeping mat type to recall it later 
    mat_type = mat_type
  ) |>
#ungrouping
  ungroup()
```

```{r}
#getting data into the average values
avg_july_dat <-july_23_dat|>
#grouping by the sample and the time
  group_by(sample, `hours after`)|>
#calculating the mean and the standard deviation from out 3 replicas at each site
  summarise(
    avg=mean(ooid), 
    sd = sd(ooid), 
    #keeping mat type to recall it later 
    mat_type = mat_type
  ) |>
#ungrouping
  ungroup()
```

## The actual plot itself
```{r}
#using new data set
ooid_plot_23_may<-avg_may_dat |>
  ggplot()+
#time is on the y axis, vs the avg ooid loss, shape is by sample
  aes(x=`hours after`, 
      y =avg, 
      color = mat_type) +
scale_color_manual(values=c('pink','darkgreen')) + 
                     
#adding error bars to our samples
  geom_errorbar(aes(
    ymin=(avg-sd),
    ymax=(avg+sd)),
    color= "darkgrey",
    width = 0.35 )+

#now adding to points so they plot on top 
  geom_point(aes(shape=sample),
             size = 3.5, show.legend = FALSE)+
    
  #labels
  labs(
    x=" ",
    # this is left blank for the cowplot 
    y=" ",
    title ='May',
    shape = "experiment loction", 
    color ="type of microbial mat"
  ) +
  #trying to edit the legend
 theme( 
   legend.position = "right"
)+
  expand_limits(y=c(0, 80))+
  theme_bw()

ooid_plot_23_may
```

```{r}
#using new data set
ooid_plot_23_feb<-avg_feb_dat |>
#  filter(`hours after` <= 125)|>
  ggplot()+
#time is on the y axis, vs the avg ooid loss, shape is by sample
  aes(x=`hours after`, 
      y =avg, 
      color = mat_type) +
scale_color_manual(values=c('pink','darkgreen')) + 
                     
#adding error bars to our samples
  geom_errorbar(aes(
    ymin=(avg-sd),
    ymax=(avg+sd)),
    color= "darkgrey",
    width = 0.35
  )+

#now adding to points so they plot on top 
  geom_point(aes(shape=sample), size = 3.5, show.legend = FALSE)+
    
  #labels
  labs(
    x=" ",
    y=" ", 
    shape = "experiment loction", 
    color ="type of microbial mat",
      title ='Febuary'
  ) +
  #trying to edit the legend
 theme( 
   legend.position = 'none'
)+
  expand_limits(y=c(0, 80))+
  theme_bw()
ooid_plot_23_feb
```

```{r}
#using new data set
ooid_plot_23_july<-avg_july_dat |>
  ggplot()+
  
#time is on the y axis, vs the avg ooid loss, shape is by sample
  aes(x=`hours after`, 
      y =avg, 
      color = mat_type) +
scale_color_manual(values=c('pink','tan','darkgreen')) + 
                     
#adding error bars to our samples
  geom_errorbar(aes(
    ymin=(avg-sd),
    ymax=(avg+sd)),
    color= "darkgrey",
    width = 0.35
  )+

#now adding to points so they plot on top 
  geom_point(aes(shape=sample),
             size = 3.5, show.legend = FALSE
              )+
    
  #labels
  labs(
    x=" ",
    y=" ", 
    shape = "experiment loction", 
    color ="type of microbial mat", 
    title ='July'
  ) +
  #trying to edit the legend
 theme( 
   legend.position = 'none'
)+
  expand_limits(y=c(0, 80))+
  theme_bw()

ooid_plot_23_july
```

## combing all three of the plots

```{r}

OP_23F<-ooid_plot_23_feb +
  labs(y="ooid coverage (%)")

OP_23M<-ooid_plot_23_may +
  labs(x="Hours after deployment")

    
# # 1. Open a pdf file
pdf("23_mat_TB_all.pdf",
    #width of pdf
    width = 7.0,
    #height of pdf
    height = 3.0)
  
# 2. plot 
plot_grid(OP_23F, OP_23M, ooid_plot_23_july,
          nrow=1) 

# Close the pdf file
dev.off() 
```

`
3.2.1 24-hr field incubation experiment results (June 2022)


```{r}
#loading the combined xcel sheet 
june_22_dat <- readxl::read_xlsx("data/ooid_data_master.xlsx", sheet = "field_june_22")

```



```{r}

#filter for our two time points (0 and 24)
H24_june_dat<-june_22_dat |>
  filter(time == c(0,24))

#the plot!!
ooid_plot_22 <- H24_june_dat|>
  ggplot() +
  aes(x=time, y =ooid, color=mat_type) +
   #adding colors for each mat_type
  scale_color_manual(values=c('lightgreen',"pink",'darkgreen')) + 
  geom_point(size=3.5, show.legend = FALSE)+
  
 #labels
  labs(
    x="time (hours)",
    y="ooid coverage (%)")+ 
  
  facet_wrap(~energy, labeller = label_both) +
  expand_limits(y=c(0, 80))+
  
  theme_bw()

ooid_plot_22

```
