---
title: "all tide"
author: "Brianna Hibner"
date: "2023-10-08"
output: html_document
---

```{r}
# load packages
library(ggplot2)
library(tidyr)
library(dplyr)
#working with time 
library(lubridate)
library(scales)
library(cowplot)

#rollmean is in the package zoo 
library(zoo)

```

## Loading the data

you have to manually go in and delete the words "logged" from un-useful columns
and re name columns (hard to do in R)

```{r}

#loading the combined xcel sheet 
tide02_II <- readxl::read_xlsx("data/372_LAC_0223.xlsx", 
    #skipping first line 
    skip=1, col_types = c("text", "guess", "numeric", "numeric", "numeric","numeric")) |>
  #dropping Nas
  drop_na()  |>
    #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-02-13 08:20:00"))


tide02_MI <- readxl::read_xlsx("data/211_LAC_0223.xlsx", 
    #skipping first line 
    skip=1,col_types = c("text", "guess", "numeric", "numeric", "numeric","numeric")) |>
   #dropping Nas
  drop_na() |>
      #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-02-13 08:20:00"))


tide02_PA <- readxl::read_xlsx("data/678_LAC_0223.xlsx", 
   #skipping first line 
  skip=1,col_types = c("text", "guess", "numeric", "numeric", "numeric","numeric")) |>
   #dropping Nas
  drop_na() |>
      #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-02-13 08:20:00"))



#loading the combined xcel sheet 
tide05_II <- readxl::read_xlsx("data/211_LAC_0523.xlsx", 
 #skipping first line                            
 skip=1)|>
   #dropping Nas
  drop_na()

tide05_MI <- readxl::read_xlsx("data/372_LAC_0523.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na()

tide05_PA <- readxl::read_xlsx("data/678_LAC_0523.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na()

#loading the combined xcel sheet 
tide07_II <- readxl::read_xlsx("data/211_LAC_0723.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na() |>
    #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-07-26 10:00:00"))

tide07_MI <- readxl::read_xlsx("data/372_LAC_0723.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na() |>
    #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-07-26 10:00:00"))

tide07_PA <- readxl::read_xlsx("data/678_LAC_0723.xlsx", 
                 #skipping first line 
                  skip=1) |>
            #dropping Nas
              drop_na() |>
  #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-07-26 10:00:00"))
  
```


##Functions for data clean and processing

noNegdepth ()   This is a function that replaces sub zero depths with 0s
the data frame entered needs to be the data$depth to work!!
returns list of values, thus pipe it to data$depth to do the old ~do si do~

GimmeMeanLine()  This is a function that gives a rolling mean for 10 data points 

```{r}
#function for replacing negative values  
noNegdepth <- function(df) {
  #returning the line of values with 0's replaced
  return(test<-ifelse(
    #if the value is less than or equal to 0 
    df <= 0,
    #replace with a 0
    0,
    #and if not, give it the old razzle dazzle (OG value of depth)
    df))
}


#function time 
GimmeMeanLine <- function(df) {
  return(ave10depth <- rollmean(df,10, fill = list(NA, NULL, NA)))
}
  

#getting the experiment time from starting line
exp_time <- function(df) {
    #calling the experiment start time (first line)
      start_time <- df[1]
      return(as.numeric(difftime(df,start_time, units='hours')))

}

#function for replacing negative values  
Clean.Depth <- function(df,cutoff) {
  #returning the line of values with 0's replaced
  return(test<-ifelse(
    #if the value is less than or equal to 0.04 
    df <= cutoff,
    #replace with a zero (no water)
    0,
    #and if not, give it the old razzle dazzle (OG value of depth)
    df))
}
```

Histrogram of values to cut ones that are"non neg"

Site PA may need to be treated differently  0.02 cm seems to be the cutoff

II and MI seem to have 0.045 cut off

```{r}
#site II
hist(tide02_II$depth,
     xlim=c(0.00,0.06),
     breaks = 100)
hist(tide02_II$depth,
     xlim=c(0.03,0.06),
     breaks = 100)

#site MI
hist(tide02_MI$depth,
     xlim=c(0.00,0.06),
     breaks = 100)
hist(tide02_MI$depth,
     xlim=c(0.04,0.06),
     breaks = 100)

#site PA
hist(tide02_PA$depth,
     xlim=c(0.00,0.06),
     breaks = 100)
hist(tide02_PA$depth,
     xlim=c(0.01,0.035),
     breaks = 100)

#may 
#site II
hist(tide05_II$depth,xlim=c(0.00,0.06),breaks = 100)
hist(tide05_II$depth,xlim=c(0.01,0.03), breaks = 100)

#site MI
hist(tide05_MI$depth,xlim=c(0.00,0.06), breaks = 100)
hist(tide05_MI$depth,xlim=c(0.02,0.04),breaks = 100)

#site PA
hist(tide05_PA$depth, xlim=c(0.00,0.06),breaks = 100)
hist(tide05_PA$depth,xlim=c(0.0,0.02), breaks = 100)

#July
#site II
hist(tide07_II$depth,xlim=c(0.00,0.06),breaks = 100)
hist(tide07_II$depth,xlim=c(0.01,0.03), breaks = 100)

#site MI
hist(tide07_MI$depth,xlim=c(0.00,0.06), breaks = 100)
hist(tide07_MI$depth, xlim=c(0.02,0.04), breaks = 100)

#site PA
hist(tide07_PA$depth,xlim=c(0.00,0.06), breaks = 100)
hist(tide07_PA$depth,xlim=c(0.0,0.02),breaks = 100)

```

##data clean (scrub a dub dub)
cutoff values are from looking at histograms above

```{r}
#febuary Tides
tide02_II$depth<- Clean.Depth(tide02_II$depth, 0.045)
tide02_MI$depth<- Clean.Depth(tide02_MI$depth, 0.045)
tide02_PA$depth<- Clean.Depth(tide02_PA$depth, 0.015)

#may tides
tide05_II$depth<- Clean.Depth(tide05_II$depth, 0.0215)
tide05_MI$depth<- Clean.Depth(tide05_MI$depth, 0.03)
tide05_PA$depth<- Clean.Depth(tide05_PA$depth, 0.01)


#july tides
tide07_II$depth<- Clean.Depth(tide07_II$depth, 0.02)
tide07_MI$depth<- Clean.Depth(tide07_MI$depth, 0.03)
tide07_PA$depth<- Clean.Depth(tide07_PA$depth, 0.01)

```

Start of experiment times

Feb
all -2/13/2023  8:37:00 AM

may
all - 5/1/2023 12:00:oo pm 
(these start at noon so they are ok ?)

July 
all - 7/26/2023  9:20:00 AM

```{r}
tide02_II$exp_time<- exp_time(tide02_II$datetime)
tide02_MI$exp_time<- exp_time(tide02_MI$datetime)
tide02_PA$exp_time<- exp_time(tide02_PA$datetime)

tide05_II$exp_time<- exp_time(tide05_II$datetime)
tide05_MI$exp_time<- exp_time(tide05_MI$datetime)
tide05_PA$exp_time<- exp_time(tide05_PA$datetime)

tide07_II$exp_time<- exp_time(tide07_II$datetime)
tide07_MI$exp_time<- exp_time(tide07_MI$datetime)
tide07_PA$exp_time<- exp_time(tide07_PA$datetime)

```


##rolling means 
```{r}
#getting rolling means for 
#feb
tide02_meanline_II<- GimmeMeanLine(tide02_II$depth)
tide02_meanline_MI<- GimmeMeanLine(tide02_MI$depth)
tide02_meanline_PA<- GimmeMeanLine(tide02_PA$depth)

#may
tide05_meanline_II<- GimmeMeanLine(tide05_II$depth)
tide05_meanline_MI<- GimmeMeanLine(tide05_MI$depth)
tide05_meanline_PA<- GimmeMeanLine(tide05_PA$depth)

#July
tide07_meanline_II<- GimmeMeanLine(tide07_II$depth)
tide07_meanline_MI<- GimmeMeanLine(tide07_MI$depth)
tide07_meanline_PA<- GimmeMeanLine(tide07_PA$depth)
 
```

##Ploting tings
```{r}
cleanDateX <-   
  #making the x nice
  scale_x_datetime(
    #null for no legend title 
    breaks = date_breaks("24 hours"),
    labels = date_format(format = " %d"))
```


#all tide plot 
```{r}

July.m.tide <-ggplot() +
#average mean line
  
     geom_line(aes(tide07_II$datetime,tide07_meanline_II), color="#CDCD00",linewidth=0.5)+
  
      geom_line(aes(tide07_MI$datetime,tide07_meanline_MI), color="#8FBC8F",linewidth=0.5)+
  
   geom_line(aes(tide07_PA$datetime,tide07_meanline_PA), color="#CD9B9B",linewidth=0.5)+
   #PA
 # cleanDateX +
   #scaling the Y axis to all reflect same range
  scale_y_continuous(limits=c(-0.025,0.43))+ 
  
  labs(x=NULL, y=NULL, title=NULL) +
  
  theme_bw() +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
      scale_x_datetime(
    #null for no legend title
    name = NULL, 
    breaks = date_breaks("1 days"),
    labels = date_format(format = "%b %d")) 

#for your viewing pleasure
July.m.tide 
  
```

#may tide
```{r}
May.m.tide <-ggplot() +
#average mean line
   #MI
   geom_line(aes(tide05_PA$datetime,tide05_meanline_PA), color="#CD9B9B", linewidth=0.5)+
     #PA
   geom_line(aes(tide05_II$datetime,tide05_meanline_II), color="#CDCD00", linewidth=0.5)+
  
  geom_line(aes(tide05_MI$datetime,tide05_meanline_MI), color="#8FBC8F", linewidth=0.5)+

  #cleanDateX +
   #scaling the Y axis to all reflect same range
  scale_y_continuous(limits=c(-0.025,0.43))+ 
  
  labs(
    x=NULL,
    y=NULL,
    title= NULL) + 
    
  theme_bw() +
  
    scale_x_datetime(
    #null for no legend title
    name = NULL, 
    breaks = date_breaks("1 days"),
    labels = date_format(format = "%b %d"))+ 
  
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

#for your viewing pleasure
May.m.tide 
```

#feb tide
```{r}
Feb.m.tide <- ggplot() +
#average mean line
  #_II
   geom_line(aes(tide02_II$datetime,tide02_meanline_II), color="#CDCD00", linewidth=0.5)+
  #MI
  geom_line(aes(tide02_MI$datetime,tide02_meanline_MI), color="#8FBC8F", linewidth=0.5)+
  #PA
   geom_line(aes(tide02_PA$datetime,tide02_meanline_PA), color="#CD9B9B", linewidth=0.5)+ 
  #cleanDateX +
   #scaling the Y axis to all reflect same range
  scale_y_continuous(limits=c(-0.025,0.43))+ 
  
  labs(
    x=NULL,
    y=NULL,
    title=NULL) +
  
    scale_x_datetime(
    #null for no legend title
    name = NULL, 
    breaks = date_breaks("1 days"),
    labels = date_format(format = "%b %d"))+ 
  
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))+
    
  theme_bw()


#for your viewing pleasure
Feb.m.tide 
```
##all together now!
```{r}

#plotting all of them 
plot_grid(Feb.m.tide,May.m.tide,July.m.tide,
          #making tide height the same
          align = "v",
          label_size = 8,
          #in three rows
          nrow=3)

ggsave("allTide_clean.pdf", width = 5, height = 3)


```

```{r}
# island interior 
max(tide02_II$depth)
max(tide05_II$depth)
max(tide07_II$depth)

#% time out of water  (exposed)
#filtering so when depth >0 (underwater) / total number of measurements 
nrow(tide02_II|> filter(depth > 0))/nrow(tide02_II) * 100
nrow(tide05_II|> filter(depth >0)) / nrow(tide05_II) *100 
nrow(tide07_II|> filter(depth > 0)) / nrow(tide07_II) *100 

```

```{r}
#MANGROVE interior
max(tide02_MI$depth)
max(tide05_MI$depth)
max(tide07_MI$depth)

#% time out of water  (exposed)
#filtering so when depth >0 (underwater) / total number of measurements 
nrow(tide02_MI|> filter(depth > 0)) / nrow(tide02_MI) * 100
nrow(tide05_MI|> filter(depth > 0)) / nrow(tide05_MI) * 100 
nrow(tide07_MI|> filter(depth > 0)) / nrow(tide07_MI) *100 
```

```{r}
# island interior 
max(tide02_PA$depth)
max(tide05_PA$depth)
max(tide07_PA$depth)

#filtering so when depth >0 (underwater) / total number of measurements 
nrow(tide02_PA|> filter(depth > 0))/ nrow(tide02_PA) * 100
nrow(tide05_PA|> filter(depth >0)) / nrow(tide05_PA) *100 
nrow(tide07_PA|> filter(depth > 0)) / nrow(tide07_PA) *100 
```

