---
title: "all tide"
author: "Brianna Hibner"
date: "2023-10-08"
output: html_document
---

```{r}
# load packages
library(ggplot2)
library(tidyr)
library(dplyr)
#working with time 
library(lubridate)
library(scales)
library(cowplot)

#rollmean is in the package zoo 
library(zoo)

```

## Loading the data

you have to manually go in and delete the words "logged" from un-useful columns
and re name columns (hard to do in R)

```{r}

#loading the combined xcel sheet 
tide02_BHSE1 <- readxl::read_xlsx("data/211_LAC_0223.xlsx", 
    #skipping first line 
    skip=1, 
    col_types = c("text", "guess", "numeric", "numeric", "numeric","numeric")) |>
  #dropping Nas
  drop_na()  |>
    #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-02-13 08:20:00"))


tide02_BHSE2 <- readxl::read_xlsx("data/372_LAC_0223.xlsx", 
                               #skipping first line 
                               skip=1,
                               col_types = c("text", "guess", "numeric", "numeric", "numeric","numeric")) |>
   #dropping Nas
  drop_na() |>
      #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-02-13 08:20:00"))


tide02_BHSE3 <- readxl::read_xlsx("data/678_LAC_0223.xlsx", 
                               #skipping first line 
                               skip=1,
                               col_types = c("text", "guess", "numeric", "numeric", "numeric","numeric")) |>
   #dropping Nas
  drop_na() |>
      #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-02-13 08:20:00"))



#loading the combined xcel sheet 
tide05_BHSE1 <- readxl::read_xlsx("data/211_LAC_0523.xlsx", 
                               #skipping first line 
                               skip=1)|>
   #dropping Nas
  drop_na()

tide05_BHSE2 <- readxl::read_xlsx("data/372_LAC_0523.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na()

tide05_BHSE3 <- readxl::read_xlsx("data/678_LAC_0523.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na()

#loading the combined xcel sheet 
tide07_BHSE1 <- readxl::read_xlsx("data/211_LAC_0723.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na() |>
    #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-07-26 10:00:00"))

tide07_BHSE2 <- readxl::read_xlsx("data/372_LAC_0723.xlsx", 
                               #skipping first line 
                               skip=1) |>
   #dropping Nas
  drop_na() |>
    #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-07-26 10:00:00"))

tide07_BHSE3 <- readxl::read_xlsx("data/678_LAC_0723.xlsx", 
                 #skipping first line 
                  skip=1) |>
            #dropping Nas
              drop_na() |>
  #7/26/2023  9:20:00 AM
   filter(datetime > ymd_hms("2023-07-26 10:00:00"))
  
```


##Functions for data clean and processing

noNegdepth ()   This is a function that replaces sub zero depths with 0s
the data frame entered needs to be the data$depth to work!!
returns list of values, thus pipe it to data$depth to do the old ~do si do~

GimmeMeanLine()  This is a function that gives a rolling mean for 10 data points 

```{r}
#function for replacing negative values  
noNegdepth <- function(df) {
  #returning the line of values with 0's replaced
  return(test<-ifelse(
    #if the value is less than or equal to 0 
    df <= 0,
    #replace with a 0
    0,
    #and if not, give it the old razzle dazzle (OG value of depth)
    df))
}


#function time 
GimmeMeanLine <- function(df) {
  return(ave10depth <- rollmean(df,10, fill = list(NA, NULL, NA)))
}
  

#getting the experiment time from starting line
exp_time <- function(df) {
    #calling the experiment start time (first line)
      start_time <- df[1]
      return(as.numeric(difftime(df,start_time, units='hours')))

}

```


##data clean (scrub a dub dub)
```{r}
#febuary tides 
tide02_BHSE1$depth<- noNegdepth(tide02_BHSE1$depth)
tide02_BHSE2$depth<- noNegdepth(tide02_BHSE2$depth)
tide02_BHSE3$depth<- noNegdepth(tide02_BHSE3$depth)

#may tides 
tide05_BHSE1$depth<- noNegdepth(tide05_BHSE1$depth)
tide05_BHSE2$depth<- noNegdepth(tide05_BHSE2$depth)
tide05_BHSE3$depth<- noNegdepth(tide05_BHSE3$depth)

#july tides
tide07_BHSE1$depth<- noNegdepth(tide07_BHSE1$depth)
tide07_BHSE2$depth<- noNegdepth(tide07_BHSE2$depth)
tide07_BHSE3$depth<- noNegdepth(tide07_BHSE3$depth)

```

##diff time??
need to:
2) find actual start times of exp (ugh) and start them there ??

Feb
all -2/13/2023  8:37:00 AM

may
all - 5/1/2023 12:00:oo pm 
(these start at noon so they are ok ?)

July 
all - 7/26/2023  9:20:00 AM

```{r}
tide02_BHSE1$exp_time<- exp_time(tide02_BHSE1$datetime)
tide02_BHSE2$exp_time<- exp_time(tide02_BHSE2$datetime)
tide02_BHSE3$exp_time<- exp_time(tide02_BHSE3$datetime)

tide05_BHSE1$exp_time<- exp_time(tide05_BHSE1$datetime)
tide05_BHSE2$exp_time<- exp_time(tide05_BHSE2$datetime)
tide05_BHSE3$exp_time<- exp_time(tide05_BHSE3$datetime)

tide07_BHSE1$exp_time<- exp_time(tide07_BHSE1$datetime)
tide07_BHSE2$exp_time<- exp_time(tide07_BHSE2$datetime)
tide07_BHSE3$exp_time<- exp_time(tide07_BHSE3$datetime)

```


##rolling means 
```{r}
#getting rolling means for 
#feb
tide02_meanline_BHSE1<- GimmeMeanLine(tide02_BHSE1$depth)
tide02_meanline_BHSE2<- GimmeMeanLine(tide02_BHSE2$depth)
tide02_meanline_BHSE3<- GimmeMeanLine(tide02_BHSE3$depth)

#may
tide05_meanline_BHSE1<- GimmeMeanLine(tide05_BHSE1$depth)
tide05_meanline_BHSE2<- GimmeMeanLine(tide05_BHSE2$depth)
tide05_meanline_BHSE3<- GimmeMeanLine(tide05_BHSE3$depth)

#July
tide07_meanline_BHSE1<- GimmeMeanLine(tide07_BHSE1$depth)
tide07_meanline_BHSE2<- GimmeMeanLine(tide07_BHSE2$depth)
tide07_meanline_BHSE3<- GimmeMeanLine(tide07_BHSE3$depth)
 
```

##Ploting tings
```{r}
cleanDateX <-   
  #making the x nice
  scale_x_datetime(
    #null for no legend title 
    breaks = date_breaks("24 hours"),
    labels = date_format(format = " %d"))
```


#all tide plot 
```{r}


July.m.tide <-ggplot() +
#average mean line
  geom_line(aes(tide07_BHSE1$datetime,tide07_meanline_BHSE1), color="#8FBC8F",linewidth=0.5)+
   #bhse3
   geom_line(aes(tide07_BHSE3$datetime,tide07_meanline_BHSE3), color="#CDCD00",linewidth=0.5)+
   #bhse2
   geom_line(aes(tide07_BHSE2$datetime,tide07_meanline_BHSE2), color="#CD9B9B",linewidth=0.5)+
  
 # cleanDateX +
   #scaling the Y axis to all reflect same range
  scale_y_continuous(limits=c(-0.025,0.43))+ 
  
  labs(
    x=NULL,
    y=NULL,
    title=NULL) +
    
  theme_bw() +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
      scale_x_datetime(
    #null for no legend title
    name = NULL, 
    breaks = date_breaks("1 days"),
    labels = date_format(format = "%b %d")) 

#for your viewing pleasure
July.m.tide 
  
```

#may tide
```{r}
May.m.tide <-ggplot() +
#average mean line
   #bhse2
   geom_line(aes(tide05_BHSE2$datetime,tide05_meanline_BHSE2), color="#CD9B9B", linewidth=0.5)+
     #bhse3
   geom_line(aes(tide05_BHSE3$datetime,tide05_meanline_BHSE3), color="#CDCD00", linewidth=0.5)+
  
  geom_line(aes(tide05_BHSE1$datetime,tide05_meanline_BHSE1), color="#8FBC8F", linewidth=0.5)+

  #cleanDateX +
   #scaling the Y axis to all reflect same range
  scale_y_continuous(limits=c(-0.025,0.43))+ 
  
  labs(
    x=NULL,
    y=NULL,
    title= NULL) + 
    
  theme_bw() +
  
    scale_x_datetime(
    #null for no legend title
    name = NULL, 
    breaks = date_breaks("1 days"),
    labels = date_format(format = "%b %d"))+ 
  
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

#for your viewing pleasure
May.m.tide 
```

#feb tide
```{r}
Feb.m.tide <- ggplot() +
#average mean line

   #bhse2
   geom_line(aes(tide02_BHSE2$datetime,tide02_meanline_BHSE2), color="#CD9B9B", linewidth=0.5)+ 
     #bhse3
   geom_line(aes(tide02_BHSE3$datetime,tide02_meanline_BHSE3), color="#CDCD00", linewidth=0.5)+
  
  geom_line(aes(tide02_BHSE1$datetime,tide02_meanline_BHSE1), color="#8FBC8F", linewidth=0.5)+
  
  #cleanDateX +
   #scaling the Y axis to all reflect same range
  scale_y_continuous(limits=c(-0.025,0.43))+ 
  
  labs(
    x=NULL,
    y=NULL,
    title=NULL) +
  
    scale_x_datetime(
    #null for no legend title
    name = NULL, 
    breaks = date_breaks("1 days"),
    labels = date_format(format = "%b %d"))+ 
  
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))+
    
  theme_bw()


#for your viewing pleasure
Feb.m.tide 
```
##all together now!
```{r}


#plotting all of them 
plot_grid(Feb.m.tide,May.m.tide,July.m.tide,
          #making tide height the same
          align = "v",
          label_size = 8,
          #in three rows
          nrow=3)

#ggsave("allTideJan07.pdf", width = 5, height = 3)


```



